\chapter{Introduction}

\section{Quantum-Resilient Communication as an Architectural Problem}

Secure communication infrastructure is entering a transition in which the
cryptographic assumptions used by deployed networks, cloud systems, and
datacenter fabrics can no longer be treated as stable design constants. The
expected arrival of cryptographically relevant quantum computers motivates the
migration from classical public-key mechanisms toward post-quantum
cryptography, while quantum key distribution and integrated quantum photonics
offer complementary approaches for generating and distributing secret material.
TODO: citation.

This transition is not only an algorithm-selection problem. A cryptographic
primitive that is secure in isolation may still be impractical if it cannot meet
the latency, throughput, determinism, power, or observability constraints of the
network in which it is deployed. Similarly, a quantum communication primitive
cannot become part of operational infrastructure unless the underlying photonic
devices, network topology, post-processing pipeline, and classical integration
points are engineered together. The central premise of this dissertation is
therefore that scalable quantum-resilient communication must be studied as an
architecture problem across device, protocol, and systems layers.

\section{Research Context and Scope}

The work in this dissertation was carried out in the context of the QUARC
project and in close collaboration with NVIDIA. The research combines academic
questions about quantum-secure communication with deployment constraints drawn
from programmable network infrastructure, especially BlueField-class data
processing units (DPUs), high-throughput datacenter communication, and
accelerated cryptographic workloads. It also reflects the Training and
Supervision Plan objective of studying the co-integration of quantum
communication technologies and post-quantum cryptography.

The scope is intentionally cross-layer. At the device layer, the dissertation
studies integrated photonic entanglement sources intended for deployable,
wavelength-division multiplexed quantum communication. At the protocol and
post-processing layer, it considers how QKD systems and post-quantum
cryptographic mechanisms can be made more practical for multi-user and
high-throughput settings. At the infrastructure layer, it evaluates how
cryptographic and security-monitoring workloads behave when mapped onto
programmable network hardware and datacenter fabrics.

Some related work is kept outside the thesis-body paper set. In particular,
ROSE-QKD is treated only as patent and title-level context unless approved
disclosure language is provided. It informs the broader research direction on
scalable QKD architecture, but it is not included as a thesis-body manuscript.

\section{Research Questions}

This dissertation is organized around the following research questions.

\begin{description}
\item[RQ1] How can post-quantum cryptographic workloads be accelerated and
deployed efficiently on programmable network infrastructure?
\item[RQ2] What architectural choices make QKD systems more scalable and
practical in multi-user communication networks?
\item[RQ3] How can integrated photonic entanglement sources be designed and
characterized for deployable, wavelength-division multiplexed QKD systems?
\item[RQ4] How do device-level, protocol-level, and infrastructure-level design
choices interact when building scalable quantum-resilient communication
systems?
\end{description}

These questions deliberately span different layers of the communication stack.
The first focuses on classical post-quantum cryptographic deployment. The
second addresses quantum communication architecture and integration. The third
connects the feasibility of quantum-secure networking to integrated photonic
source design. The fourth binds the dissertation together by asking how choices
made at one layer constrain or enable the others.

\section{Research Pillars and Contributions}

The dissertation advances this theme through three connected research pillars.

\medskip\noindent\textbf{High-performance post-quantum cryptography.}
The first pillar investigates how post-quantum cryptographic algorithms can be
executed efficiently on heterogeneous DPU platforms. The work studies
CPU-GPU-DPU workload partitioning, batching behavior, latency and throughput
tradeoffs, and the conditions under which acceleration is useful for
datacenter-scale deployment.

\medskip\noindent\textbf{Scalable and topology-aware QKD architectures.}
The second pillar studies QKD as a deployable network architecture rather than
only as a point-to-point protocol. It considers receiver placement, topology,
post-processing, authentication, and integration with classical infrastructure.
Patent-sensitive QKD architecture work is referenced only at the title/status
level in this dissertation.

\medskip\noindent\textbf{Integrated photonic entanglement sources.}
The third pillar addresses the device-level foundations needed for practical
quantum-secure networking. It focuses on dual-coupled microring resonator
sources, spectral stability, thermal tuning, Q-factor engineering, and
compatibility with telecom wavelength-division multiplexing.

Together, these pillars support the thesis claim that quantum-resilient
communication systems must be designed and evaluated across layers. The
dissertation contributes paper-based studies of integrated photonic source
behavior, QKD post-processing on programmable network hardware, DPU-accelerated
post-quantum cryptography, post-quantum secured edge AI pipelines,
ordering-related risks in RDMA over optical fabrics, and observability for
confidential AI infrastructure.

\section{Dissertation Organization}

This is a compilation-style dissertation. The main body is organized around
published or accepted manuscripts, with thesis-formatted wrapper sections that
explain each paper's role in the broader argument. Each paper-bearing chapter
introduces the chapter motivation, relates the included work to the research
questions, records publication context and author contribution, includes or
points to the manuscript, and then discusses the paper's implications for the
next layer of the thesis.

Chapter 2 provides the background and unifying concepts needed to read the
paper-centric chapters. Chapter 3 begins at the photonic device layer with
integrated quantum sources. Chapter 4 moves from quantum key generation toward
QKD post-processing on programmable network hardware. Chapters 5 and 6 study
post-quantum cryptographic acceleration and deployment in DPU-based systems.
Chapters 7 and 8 examine datacenter fabric security and confidential AI
observability. Chapter 9 synthesizes the cross-layer lessons, limitations, and
future outlook.
