\chapter{Background and Unifying Concepts}
\label{ch:background}

\section{Post-Quantum Cryptography and Deployment Constraints}

Post-quantum cryptography (PQC) replaces public-key mechanisms vulnerable to
large-scale quantum attacks with algorithms believed to resist both classical
and quantum adversaries. For networked systems, however, the migration to PQC is
not finished when an algorithm is selected. Key encapsulation mechanisms,
digital signatures, certificates, handshakes, and identity-provisioning systems
must all meet operational latency and throughput requirements while preserving
security margins and interoperability. TODO: citation.

The included PQC work studies this deployment problem through workloads such as
ML-KEM, ML-DSA, and HQC. These algorithms differ in computational structure,
memory behavior, message size, and suitability for batching. As a result, the
right execution strategy depends on both the cryptographic operation and the
system context in which it runs. A mechanism that performs well for a single
operation on a host CPU may behave differently when thousands of operations must
be processed under datacenter provisioning, service-mesh, or secure-channel
constraints.

The central deployment questions are therefore architectural: where should
cryptographic work run, how should requests be batched, how much latency can be
tolerated, what resources become bottlenecks, and how deterministic is the
result under load? These questions motivate the use of programmable network
infrastructure as a recurring platform in this dissertation.

\section{Programmable Network Infrastructure for Security Offload}

SmartNICs and DPUs extend the network interface from a passive I/O endpoint into
a programmable execution environment. In BlueField-class systems, Arm CPU cores,
network accelerators, memory resources, and in some settings GPU-connected
execution paths can be used to move selected work away from the host. This
creates an opportunity to place cryptographic operations, packet processing,
storage or image I/O, telemetry, and policy enforcement closer to the network
boundary.

Offload is not automatically beneficial. It introduces its own costs: data
movement, queueing, synchronization, resource contention, and additional
software complexity. The useful question is not whether a DPU is faster in
isolation, but which part of a security pipeline should be moved, under which
batching and traffic conditions, and with which visibility into system state.
This dissertation uses DPUs both as accelerators for post-quantum workloads and
as observation points for security-sensitive infrastructure.

\section{QKD Networks and Scalable Architectures}

Quantum key distribution (QKD) provides a different approach to secure key
establishment by using quantum states and classical post-processing to generate
shared secret keys. A deployable QKD system is more than a quantum channel. It
requires source and detector hardware, synchronization, sifting, error
reconciliation, privacy amplification, authentication, key management, and
interfaces to classical applications.

The scalability challenge appears when QKD moves from a laboratory link to a
multi-user network. Topology, receiver placement, trusted-node assumptions,
post-processing throughput, and classical integration all influence whether a
QKD architecture can grow beyond isolated point-to-point demonstrations. This
dissertation treats scalable QKD architecture as part of the broader
quantum-resilient communication problem, while patent-sensitive concepts such as
ROSE-QKD remain limited to title-level context unless approved disclosure text
is supplied.

\section{Integrated Photonic Sources for Deployable QKD}

Integrated photonics addresses the device-level side of deployable quantum
communication. For entanglement-based QKD, source stability, spectral
engineering, tunability, quality factor, loss, and compatibility with telecom
wavelength grids all affect whether a photonic device can support practical
network operation. Device behavior therefore has system-level consequences:
source drift, resonance spacing, and thermal response influence how easily a
quantum source can be controlled, multiplexed, and integrated with classical
network infrastructure.

The photonic-source work in this dissertation focuses on dual-coupled microring
resonators as candidate sources for frequency-bin entanglement. The relevant
background includes free spectral range behavior, thermal tuning, Q-factor
engineering, lithium niobate and silicon nitride platform tradeoffs, and
wavelength-division multiplexing compatibility. These topics connect the device
chapters to the larger architectural question of how quantum communication
components can become deployable network building blocks.

\section{Cross-Layer Architecture Evaluation}

The common method across the dissertation is to identify a security or
deployment assumption at one layer, then evaluate how it changes when mapped
onto another layer. For PQC, this means studying how algorithmic costs interact
with DPU execution, batching, and datacenter-scale workloads. For QKD, it means
connecting quantum key generation and post-processing to programmable network
hardware and classical integration points. For integrated photonics, it means
relating device-level spectral behavior to network-level deployability. For
datacenter security and confidential AI, it means examining what can still be
observed or protected when payload access, ordering assumptions, or topology
assumptions are constrained.

This cross-layer perspective is the reason the dissertation combines topics
that might otherwise appear separate: quantum photonic devices, QKD
post-processing, post-quantum cryptographic acceleration, RDMA and optical
fabrics, and confidential AI observability. The unifying question is how secure
communication systems can be made scalable and deployable when the relevant
constraints are distributed across the physical, cryptographic, network, and
application layers.
