\chapter{Integrated Quantum Sources for Quantum Networks}
\label{ch:quantum-sources}

\section{From Photonic Devices to Network Components}

Quantum-secured networks depend on physical devices whose behavior must remain predictable after they leave a controlled laboratory setting. Entangled-photon sources are one such device class: they determine the spectral structure, brightness, purity, and long-term stability of the quantum states that later protocols consume. If these source properties drift with fabrication tolerances, thermal changes, or uncontrolled mode interactions, then the network layer inherits an unstable security and performance foundation.

This chapter therefore begins the thesis body at the photonic-device layer. It studies dual-coupled micro-ring resonators as integrated sources for frequency-bin entangled photons. The common thesis-level question is not only whether such a device can generate useful quantum states, but whether it can be designed, tuned, and modeled in a way that makes it suitable for larger quantum--classical network systems.

The dual-coupled architecture adds an auxiliary resonator to the main photon-pair generation ring. This additional degree of freedom can be used to reshape the spectral response of the source, compensate for nonidealities, and expose a control interface between device physics and network-facing metrics such as channel spacing, heralded purity, and Schmidt number. In this sense, the chapter treats source engineering as the first cross-layer problem of the dissertation: a physical-layer design choice must be made legible to the network and security layers built above it.

\section{Mapping Device Physics to Network Security}

The work in this chapter is written around the following mapping between low-level device behavior and the network requirements introduced in Chapter~\ref{ch:background}:

\begin{itemize}
  \item It addresses how quantum-secure networking primitives can be made compatible with deployed optical infrastructure. The focus is on C-band, wavelength-division-multiplexed operation and on resonator behavior that can be related to telecom channel plans \cite{Kumar2015,Mao2018}.
  \item It addresses how low-level device parameters become system-level security constraints. Spectral purity, free spectral range stability, and the Schmidt number are device-derived quantities, but they influence interference visibility, quantum-bit error rate, and the practical budget available for coexistence with classical traffic \cite{Grebenchukov2023}.
  \item It introduces the thesis pattern of cross-layer abstraction. Rather than treating a quantum source as a black box, the chapter asks which control variables and models are needed before a source can be used as a predictable component in a quantum network.
\end{itemize}

This chapter also sets up the transition to programmable network infrastructure in Chapter~\ref{ch:qkd-dpu}. The quantum source provides the raw optical resource, while later chapters ask how the resulting key material, cryptographic work, and observability tasks can be handled by DPUs, SmartNICs, and datacenter network systems.

\section{Contribution and Articles}

This chapter is built around two contributions that, taken together, move the dual-coupled microring (DCM) source from a promising integrated-photonics component toward a predictable network-engineering object. The author contributed to connecting the device studies to network requirements, the simulation-driven evaluation, and the thesis-level system interpretation; the experimental device work, modeling, and supervision were collaborative efforts with the co-authors.

The first contribution (P1), \enquote{Simulation-Driven Optimization of Dual-Coupled Micro-Ring Entanglement Sources for Quantum Networks}, submitted to the 26th International Conference on Transparent Optical Networks (ICTON 2026), establishes the theoretical and modeling foundation for the chapter. It presents a coupled-mode-theory digital twin for the DCM source, relating auxiliary-cavity detuning, inter-ring coupling, and dispersion to photon-pair purity, brightness, and Schmidt number. By mapping low-level physical parameters to network-facing metrics, it defines the optimization space required to engineer quantum sources for specific network applications.

The second contribution (P2), \enquote{Free Spectral Range Stability under Linear Thermal Tuning in a Dual-Coupled Microring Resonator}, presented at the 2026 International Conference on Science and Electrical Engineering (ICSEE), provides the experimental validation of the device's stability. While the first paper demonstrated how the source can be optimized in simulation, this work grounds those findings in physical reality. Its key role is to show that thermal tuning of the auxiliary resonator---a necessary control mechanism---preserves the predictable free spectral range behavior required for channelized quantum network operation.

Together, the two papers establish that source behavior can be modeled, optimized, and physically stabilized in terms that later network layers can consume. This matters because quantum-secure communication is often limited by interfaces between layers: physical devices expose imperfect states; protocols assume idealized state properties; network planners allocate wavelengths and power budgets; and operators need stable, remotely configurable behavior. The resulting thesis claim is deliberately bounded. The chapter does not demonstrate a complete deployed quantum network, nor does it by itself close the practical-security gap of QKD systems; it establishes a modelable, stabilizable source as the foundation on which the rest of the thesis builds.

Two limitations should remain visible in the final dissertation. First, predictive source models must be continuously validated against fabricated devices across process variation and environmental drift; a digital twin that is not tied back to measurement risks becoming another idealized model. Second, source optimization is only one part of system security. Higher purity and stable channel spacing can improve the assumptions available to entanglement-based protocols, but coexistence noise, detector behavior, post-processing, authentication, and operational monitoring still determine whether a practical system remains secure \cite{Xu2020}. These limitations motivate the rest of the thesis, which follows quantum-secure networking beyond the photonic source and into programmable post-processing, cryptographic acceleration, datacenter transport behavior, and confidential-AI observability. The immediate next step, in Chapter~\ref{ch:qkd-dpu}, moves from generating quantum-secure optical resources to processing the classical information that such systems produce on programmable network hardware.

\includedpaper{P1}{Simulation-Driven Optimization of Dual-Coupled Micro-Ring Entanglement Sources for Quantum Networks}{meyer2026icton}{source-materials/papers/camera-ready/dcm-entanglement-sources-quantum-networks-camera-ready.pdf}

\includedpaper{P2}{Free Spectral Range Stability under Linear Thermal Tuning in a Dual-Coupled Microring Resonator}{meyer2026icsee}{source-materials/papers/camera-ready/fsr-stability-dual-coupled-microring-camera-ready.pdf}
