\chapter*{Summary}
\addcontentsline{toc}{chapter}{Summary}

\noindent\textbf{Cross-Layer Network Security in the Post-Quantum Era:
From Integrated Quantum Sources to Datacenter-Scale Cryptography}

\vspace{1.5em}

The security of digital communication rests on mathematical problems that are
easy to pose but, for classical computers, practically impossible to solve. The
prospect of a cryptographically relevant quantum computer (CRQC) undermines this
foundation: once such a machine exists, the public-key cryptography that protects
nearly all modern infrastructure can be broken. Although the precise arrival of
``Q-Day'' is uncertain, the threat is already operational through
``Harvest Now, Decrypt Later'' attacks, in which adversaries store encrypted
traffic today to decrypt it once the hardware matures. Mitigating this threat
is not a single algorithm-selection problem but a tiered, heterogeneous
re-engineering of communication systems. This shift is driven by fundamental
trade-offs between security guarantees, deployment complexity, and performance:
computationally secure post-quantum cryptography (PQC) scales to protect
high-volume datacenter traffic but relies on mathematical assumptions, whereas
quantum key distribution (QKD) offers unconditional, physics-based security at
the cost of specialized optical hardware and dedicated infrastructure. A
comprehensive defense must therefore integrate both paradigms.

The central premise of this dissertation is that scalable quantum-resilient
communication is fundamentally an integration challenge governed by the boundaries
between the device, protocol, and infrastructure layers. A cryptographic primitive that is
secure in isolation may still be impractical if it cannot meet the latency,
throughput, determinism, power, or observability constraints of the network in
which it runs; likewise, a quantum primitive becomes operational infrastructure
only when its photonic devices, post-processing pipeline, and classical
integration points are engineered together. The work was carried out within the
QUARC doctoral network as an industrial doctorate embedded with NVIDIA's
networking organization.

The thesis is organized as a compilation of peer-reviewed publications grouped
into three parts. Part~I treats QKD as a deployable network capability: it
presents the design and characterization of integrated photonic entanglement
sources for wavelength-division-multiplexed QKD, and the offload of QKD
post-processing onto data processing units (DPUs). Part~II investigates how PQC
algorithms can be executed efficiently on heterogeneous CPU--GPU--DPU platforms
and applied to secure edge AI pipelines. Part~III examines the programmable
datacenter substrate on which both mechanisms ultimately run, addressing the
ordering- and timing-related risks of RDMA over high-throughput optical fabrics
and cross-layer observability and intrusion detection for confidential AI
infrastructure.

Taken together, these contributions show that device-level, protocol-level, and
infrastructure-level design choices are tightly coupled: decisions made at one
layer constrain or enable the others. The synthesis argues that
quantum-resilient communication is achievable with today's and near-future
technology, but only by actively engineering the integration points where cryptography,
photonics, and programmable network hardware meet, rather than treating them as
independent disciplines.
